---
title: "P8105 HW2"
author: Kaitlyn Wang
output: html_document
---

```{r}
library(tidyverse)
library(tidyr)
library(dplyr)
```

## Problem 1

Read and clean the Mr. Trash Wheel sheet:
```{r}
trash_wheel = 
  readxl::read_excel("./data/TWC.xlsx", sheet = "Mr. Trash Wheel", range = "A2:N535") %>% 
  janitor::clean_names() %>% 
  filter(dumpster != "NA") %>% 
  mutate(sports_balls = round(sports_balls, digits = 0))
```

Read and clean precipitation data for 2018 and 2019.
```{r}
prcp2018 = readxl::read_excel("./data/TWC.xlsx", sheet = "2018 Precipitation", skip = 1) %>%
  janitor::clean_names() %>% 
  drop_na() %>% 
  mutate(year = 2018)

prcp2019 = readxl::read_excel("./data/TWC.xlsx", sheet = "2019 Precipitation", skip = 1) %>%
  janitor::clean_names() %>% 
  drop_na() %>% 
  mutate(year = 2019)

prcp = full_join(prcp2018, prcp2019) %>% 
  mutate(month = month.name[month])
```

The Mr. Trash Wheel dataset contains `r nrow(trash_wheel)` observations in `r ncol(trash_wheel)` categories. For example, year, months, and multiples trash categories.

The precipitation dataset contains `r nrow(prcp)` rows and `r ncol(prcp)` columns, including months, precipitation and year. The total precipitation in 2018 is `r sum(pull(prcp2018, total))` and the median number of sports balls in a dumpster in 2019 is `r median(pull(trash_wheel, sports_balls))`. 

## Problem 2

First, clean the data in pols-month.csv. 
```{r}
pols_month = read_csv("./data/datasets/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, into = c("year", "month", "day")) %>% 
  mutate(month = as.numeric(month), year = as.numeric(year)) %>% 
  mutate(month = month.name[month]) %>% 
  mutate(president = ifelse(prez_gop == 0, "dem", "gop")) %>% 
  select(-c(prez_dem,prez_gop,day))
```

Second, clean the data in snp.csv.
```{r}
snp = read_csv("./data/datasets/snp.csv") %>% 
  separate(date, into = c("month", "day", "year")) %>% 
  mutate(month = as.numeric(month), year = as.numeric(year)) %>%
  mutate(year = ifelse(year < 50, year + 2000, year + 1900)) %>% 
  arrange(year, month) %>% 
  mutate(month = month.name[month]) %>% 
  relocate(year, month)
```

Third, tidy the unemployment data so that it can be merged with the previous datasets.
```{r}
unemploy = read_csv("./data/datasets/unemployment.csv") %>%
  rename(year = Year) %>%
  pivot_longer(
    Jan:Dec,
    names_to = "month",
    values_to = "pct_unemployed"
    ) %>% 
  mutate(month = match(month, month.abb)) %>% 
  mutate(month = month.name[month])
```

Join the datasets by merging snp into pols, and merging unemployment into the result.
```{r}
list1 = left_join(pols_month, snp)
list2 = left_join(list1, unemploy)
```

The dataset contains `r nrow(list2)` rows ranging from year 1947 to year 2015; and `r ncol(list2)` columns with information including whether the president is republican or democrat and the number of senator, governor and representative in each party in different time. 


## Problem 3

Load and tidy the data.
```{r}
names_raw = read_csv("./data/Popular_Baby_Names.csv") %>%
  janitor::clean_names() %>% 
  mutate(childs_first_name = tolower(childs_first_name)) %>% 
  distinct()

unique(pull(names_raw, ethnicity))
```

```{r}
names = 
  names_raw %>% 
  mutate(ethnicity = replace(ethnicity, ethnicity == "WHITE NON HISP", "WHITE NON HISPANIC")) %>% 
  mutate(ethnicity = replace(ethnicity, ethnicity == "ASIAN AND PACI","ASIAN AND PACIFIC ISLANDER")) %>% 
  mutate(ethnicity = replace(ethnicity, ethnicity == "BLACK NON HISP", "BLACK NON HISPANIC"))
```

Produce a well-structured, reader-friendly table showing the rank in popularity of the name “Olivia” as a female baby name over time; this should have rows for ethnicities and columns for year. 
```{r}
olivia = 
  names %>% 
  filter(childs_first_name == "olivia") %>% 
  select(year_of_birth, ethnicity, rank) %>% 
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "rank"
  ) %>% 
  relocate(ethnicity, "2011", "2012", "2013", "2014", "2015", "2016")
```

Produce a similar table showing the most popular name among male children over time.
```{r}
male_pop = 
  names %>% 
  filter(gender == "MALE") %>% 
  group_by(childs_first_name) %>% 
  summarise(total = sum(count)) %>% 
  arrange(desc(total))

ethan = 
  names %>% 
  filter(childs_first_name == "ethan") %>% 
  select(year_of_birth, ethnicity, rank) %>% 
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "rank"
  ) %>% 
  relocate(ethnicity, "2011", "2012", "2013", "2014", "2015", "2016")
```

Finally, for male, white non-hispanic children born in 2016, produce a scatter plot showing the number of children with a name (y axis) against the rank in popularity of that name (x axis).
```{r}
white_2016 = 
  names %>% 
  filter(gender == "MALE", year_of_birth == "2016", ethnicity == "WHITE NON HISPANIC")

ggplot(white_2016, aes(x = rank, y = count)) + geom_point() +
  labs(
    title = "Number of Each Name vs. Rank in Popularity",
    x = "Rank in Popularity",
    y = "Number of Each Name"
  )
```

